### 2021-02-02 conclusion.tex
conclusions and 15!

\begin{document}
\section{Related Work and Conclusion}
\subsubsection*{Deadlock freedom and progress}
Deadlock freedom and progress are well studied properties in the $\pi$-calculus.
For the standard typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set. These abstract tags were later simplified to natural numbers. Pairs of obligations and capabilities were used in the type system for deadlock freedom~\cite{kobayashi06}, which allowed more $\pi$-calculus processes to be typed. Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later simplified them to a single priority for linear $\pi$-calculus~\cite{padovani14}.  As discussed by the author, by combining the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12,dardha14beat,dardha16} and the priority-based linear $\pi$-calculus, we can obtain deadlock freedom for the $\pi$-calculus with session types.

For the session-typed $\pi$-calculus, the foundational work by Dezani~\etal~\cite{dezani-ciancaglinimostrous06} guarantees progress by allowing only one active session at a time. Dezani later~\cite{dezani-ciancagliniliguoro09progress} introduces a partial order on channels in line with Kobayashi's work~\cite{kobayashi98}. Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser} used to provide a missing counterpart to a process, thus guaranteeing deadlock freedom.
Carbone~\etal~\cite{carbonedardha14} studied the use of catalysers and showed that progress is a compositional form of lock-freedom. In the same lines of Padovani~\cite{padovani14}, Carbone~\etal~\cite{carbonedardha14} show that progress can be lifted to session types via the encoding of session types to linear types and Kobayashi's obligations/capabilities. Vieira and Vasconcelos~\cite{vieiravasconcelos13} use single priorities and an abstract partial order to guarantee deadlock freedom in a session-typed $\pi$-calculus.

Multiparty Session Types (MPST) \cite{hondayoshida08} on the other hand guarantee deadlock freedom within a single session, but more advanced techniques are required for session interleaving. Bettini~\etal~\cite{bettinicoppo08} follow a similar technique to Kobayashi for multiparty session types. The main
difference with our work is that we associate priorities to communication actions, and not to channels. Carbone and Montesi~\cite{carbonemontesi13} explore and combine MPST and choreographies to obtain a formalism that satisfies deadlock freedom by design.

@@

\subsubsection*{Ties with logic}
The correspondence between logic and types lays the foundation for functional programming~\cite{wadler15}. Since its inception by Girard~\cite{girard87}, linear logic has been a candidate for a foundational correspondence for concurrent programs. A~correspondence with linear $\pi$-calculus was established early on by Abramsky~\cite{abramsky94,bellinscott94}. A~correspondence between session-typed $\pi$-calculus and dual intuitionistic linear logic was developed by Caires and Pfenning~\cite[$\pi\text{DILL}$]{cairespfenning10}, and with classical linear logic by Wadler~\cite[CP]{wadler15}, both guaranteeing deadlock freedom as a result of their connection to logic. Dardha and Gay~\cite[PCP]{dardhagay18} integrate Kobayashi's and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, creating the first calculus which combines priorities and strong ties with logic. Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and show that CLL corresponds to a subsystem of Kobayashi's where restriction is applied once. Balzer~\etal~\cite[$\text{SILL}_S$]{balzerpfenning17} introduce shared state, which breaks deadlock freedom. They later restore deadlock freedom using priorities~\cite[$\text{SILL}_{S+}$]{balzertoninho19}.
Carbone~\etal~\cite{carbonelindley16} give a logical view of MPST with a generalised duality, called coherence.

\subsubsection*{Conclusion and future work}
We presented Priority GV, a~session-typed functional language which uses priorities to ensure deadlock freedom, and showed its relation to Priority CP~\cite{dardhagay18} via an operational correspondence.

Our formalism so far only captures the core of GV. In future work, we plan to explore: recursion in PGV, by integrating the works of Lindley and Morris~\cite{lindleymorris16} and Padovani and Novara~\cite{padovaninovara15}; a~translation from Priority GV to Priority CP; and sharing, following Balzer and Pfenning~\cite{balzerpfenning17}.
\end{document}

>>>>

\begin{document}
\section{Related Work and Conclusion}
\subsubsection*{Deadlock freedom and progress}
%Deadlock freedom and progress are well studied properties in the $\pi$-calculus.
For the standard typed $\pi$-calculus, one line of work is Kobayashi's approach to deadlock freedom~\cite{kobayashi98}, where priorities are abstract tags defined over a partially ordered set. These tags were later simplified to pairs of natural numbers, called obligations and capabilities~\cite{kobayashi06}, which allowed more $\pi$-calculus processes to be typed. Padovani~\cite{padovani13} adapted the obligation and capability pairs to session types, and later simplified them to a single priority for the linear $\pi$-calculus~\cite{padovani14}.  Furthermore, by exploiting the encoding of session types into linear types~\cite{kobayashi07,dardhagiachino12}
%,dardha14beat,dardha16 %
and the priority-based linear $\pi$-calculus,
we can obtain deadlock freedom for the $\pi$-calculus with session types.

For the session-typed $\pi$-calculus, Dezani~\etal~\cite{dezani-ciancaglinimostrous06} guarantee progress by allowing only one active session at a time. Dezani later~\cite{dezani-ciancagliniliguoro09progress} introduces a partial order on channels similar to Kobayashi~\cite{kobayashi98}. Carbone and Debois~\cite{carbonedebois10} define progress for session typed $\pi$-calculus in terms of a \emph{catalyser}, which provides a missing counterpart to a process, thus guaranteeing deadlock freedom.
Carbone~\etal~\cite{carbonedardha14} use such catalysers to show that progress is a compositional form of lock-freedom and that it can be lifted to session types via the encoding of session types to linear types. Vieira and Vasconcelos~\cite{vieiravasconcelos13} use single priorities and an abstract partial order to guarantee deadlock freedom in a session-typed $\pi$-calculus.
%
Multiparty Session Types (MPST) \cite{hondayoshida08} guarantee deadlock freedom \emph{within a single} session, but not for session interleaving.
Consequently, several techniques for deadlock freedom in MPST have been defined \cite{bettinicoppo08,carbonemontesi13}.
%Bettini~\etal~\cite{bettinicoppo08} follow a similar technique to Kobayashi for multiparty session types. The main
%difference with our work is that we associate priorities to communication actions, and not to channels. Carbone and Montesi~\cite{carbonemontesi13} explore and combine MPST and choreographies to obtain a formalism that satisfies deadlock freedom by design.

@@

\subsubsection*{Ties with logic}
The correspondence between logic and types lays the foundation for functional programming~\cite{wadler15}. Since its inception by Girard~\cite{girard87}, linear logic has been a candidate for a foundational correspondence for concurrent programs. A~correspondence with linear $\pi$-calculus was established early on by Abramsky~\cite{abramsky94,bellinscott94}. A~correspondence between session-typed $\pi$-calculus and dual intuitionistic linear logic was developed by Caires and Pfenning~\cite[$\pi\text{DILL}$]{cairespfenning10}, and with classical linear logic by Wadler~\cite[CP]{wadler15}, both guaranteeing deadlock freedom as a result of their connection to logic. Dardha and Gay~\cite[PCP]{dardhagay18} integrate Kobayashi and Padovani's work on priorities~\cite{kobayashi06,padovani14} with CP, creating the first calculus which combines priorities and strong ties with logic. Dardha and P\'{e}rez~\cite{dardhaperez15} compare Kobayashi-style typing and CLL typing, and show that CLL corresponds to a subsystem of Kobayashi's where restriction is applied once. Balzer~\etal~\cite[$\text{SILL}_S$]{balzerpfenning17} introduce shared state, which breaks deadlock freedom. They later restore deadlock freedom using priorities~\cite[$\text{SILL}_{S+}$]{balzertoninho19}.
Carbone~\etal~\cite{carbonelindley16} give a logical view of MPST with a generalised duality, called coherence.

\subsubsection*{Conclusion and future work}
We answered our research question by presenting Priority GV, a~session-typed functional language which allows cyclic communication structures and uses priorities to ensure deadlock freedom. We showed its relation to Priority CP~\cite{dardhagay18} via an operational correspondence.

Our formalism so far only captures the core of GV. In future work, we plan to explore: recursion in PGV, by integrating the works of Lindley and Morris~\cite{lindleymorris16} and Padovani and Novara~\cite{padovaninovara15}; %a~translation from Priority GV to Priority CP;
and sharing, following Balzer and Pfenning~\cite{balzerpfenning17}.

\subsubsection*{Acknowledgement} The authors would like to thank Simon Fowler and Philip Wadler for they useful and insightful feedback.

\end{document}
